Chapter 2 Figures

### 2019-12-18 doc/figureFiles/figures.Rmd
removed facet ~sex, replace shape = sex. version sent to bret

# Option 1, strain averages, sexes matched

@@

#use the MLH1 / cell level data to calq the error bars
strain.table.HQ <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Nmice), 3)
    )

@@

```{r strain.table.plot.v2, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_Bv2_strain.male <- ggplot(strain.table.HQ[strain.table.HQ$sex == "male",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
  ylab("")+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3)+
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #Dom
  annotate("segment", x=.75, xend=3.6, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=3.9, xend=6.15, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=6.5, xend=8.1, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(2,5,7.15), #list of all x points 
           y = c(18.5,18.5, 18.5),

# 2
Fig1_Bv2_strain.female <- ggplot(strain.table.HQ[strain.table.HQ$sex == "female",], aes(y=mean_co, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co-se, ymax=mean_co+se), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
#mol
 annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
    #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
           y = c(18.5,18.5,18.5),
           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

```{r Strain.fig.merge.v2, echo=FALSE}

fig1_B_v2  <- ggdraw() +
  draw_plot(Fig1_Bv2_strain.female, x=0, y = 0, width=.5, height=.97) +
  draw_plot(Fig1_Bv2_strain.male, x=.45, y=0, width=.5, height=.97) +

  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .45), y = c(1, 1) )
fig1_B_v2
```

Molf female data point is missing because there is only 1 female, so NAs were generated.
Mol text should be moved back above the points.

## option 2, strain averages, more strains

@@

#use the MLH1 / cell level data to calq the error bars
strain.table.HQ.backup <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Nmice), 3)
    )

@@

```{r Strain.fig.merge, echo=FALSE}
fig1_B_v1  <- ggdraw() +
  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.5, height=.97) +
  draw_plot(Fig1_B_strain.male, x=.45, y=0, width=.5, height=.97) +

@@

>>>>

# Option 1, strain averages, sexes matched, cells pooled

@@

#step-wise, 1)mouse-level, 2)averages of mouse-levels

#Pooling all cells for a category
strain.table.HQ <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)
    )

@@

```{r strain.table.plot.v2, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot
#remove facet by sex

Fig1_Bv2_strain <- ggplot(strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape = sex))+
  #facet_grid(~sex, scales = "free", space="free")+ 
  ylab("Average MLH1 Foci per Cell")+
  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+

   geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(19.5, 33))+
 scale_color_manual(values=colors_of_strains)+ 
  #legend.position="none",
  theme(legend.position="bottom",legend.box = "horizontal",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+

#remove y labfor male 
  #Dom
  annotate("segment", x=.75, xend=3.6, y=21, yend=21, colour="black", size=1)+
  #Musc
  annotate("segment", x=3.9, xend=6.15, y=21, yend=21, colour="black", size=1)+
  #mol
  annotate("segment", x=6.5, xend=8.1, y=21, yend=21, colour="black", size=1)+
  #label
  annotate("text", x = c(2.2, 4.5, 7), #list of all x points 
           y = c(20.2,20.2,20.2),
   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

#this can pull just the lengend object
#grobs <- ggplotGrob(Fig1_B_strain.male)$grobs
#legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]
```

```{r Strain.fig.merge.v2, echo=FALSE}
Fig1_Bv2_strain
#save a tif or png file
ggsave("~./MLH1repo/doc/figureFiles/MLH1Fig1.png")
```

Error bars are 2 SE,

S.E. = sd( all cells) / n.cells

## Option 2, strain averages, more strains, pooled cells

@@

mouse.table.HQ.backup <- ddply(fig.data, c("mouse"), summarise,
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Ncells), 3) #mean is pooled cells, so sd should be pooled
    )
#need to add strain and sex
mouse.table.HQ.backup <- add_strain(mouse.table.HQ.backup)
mouse.table.HQ.backup <- add_sex(mouse.table.HQ.backup)

strain.table.from.mouse.HQ.backup <- ddply(mouse.table.HQ.backup, c("strain","sex"), summarise,
                  Nmice  = length(mouse),
                  mean_co.of.mice = as.numeric(format(round(  mean(mean_co), 3 ), nsmall=3) ),
                  cV.of.mice = cv(mean_co, na.rm = TRUE),
                   var.of.mice = as.numeric(format(round(   var(mean_co, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                  sd.of.mice   = round(sd(mean_co, na.rm = TRUE), 3),
                  se.of.mice   = round(sd.of.mice / sqrt(Nmice), 3) #mean is pooled cells, so sd should be pooled
    )

#use the MLH1 / cell level data to calq the error bars
strain.table.HQ.backup <- ddply(fig.data, c("strain", "sex"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), #this turns to chr
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                        se   = round(sd / sqrt(Ncells), 3) #mean is pooled cells, so sd should be pooled
    )

@@

```{r Strain.fig.merge, echo=FALSE}
fig1_B_v1  <- ggdraw() +
  draw_plot(Fig1_B_strain.female, x=0, y = 0, width=.45, height=.97) +
  draw_plot(Fig1_B_strain.male, x=.5, y=0, width=.45, height=.97) +

@@

## Option 2.B strain averages, mouse observations

```{r strain.table.plot.mouse.level, echo=FALSE, warning=FALSE}
#use the MLH1 / cell level data to calq the error bars
#strain table ggplot

Fig1_C_strain.male <- ggplot(strain.table.from.mouse.HQ.backup[strain.table.from.mouse.HQ.backup$sex == "male",], aes(y=mean_co.of.mice, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ 
  ylab("")+
  geom_point(size=3)+
  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2) , ymax=mean_co.of.mice+(se.of.mice*2) ), width=.3)+
  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10), #list of all x points 
           y = c(18.5,18.5,33),

# 2
Fig1_C_strain.female <- ggplot(strain.table.from.mouse.HQ.backup[strain.table.from.mouse.HQ.backup$sex == "female",], aes(y=mean_co.of.mice, x=strain, color=strain))+facet_grid(~sex, scales = "free", space="free")+ ylab("Average MLH1 Foci per Cell")+
  geom_point(size=3)+
  ylim(c(18, 35))+
  geom_errorbar(aes(ymin=mean_co.of.mice-(se.of.mice*2), ymax=mean_co.of.mice+(se.of.mice*2)  ), width=.3) +
  scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  #segments
  annotate("segment", x=.75, xend=3.5,y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x=4, xend=6.5, y=19.5, yend=19.5, colour="black", size=1)+
#mol
 annotate("segment", x=7, xend=8.5, y=19.5, yend=19.5, colour="black", size=1)+
    #Cast
#  annotate("segment", x=4, xend=2, y=18, yend=18, colour="black", size=1)
#Labels
 annotate("text", x = c(2.25,5.6,6.5), #list of all x points 
           y = c(18.5,18.5,33),
           label = c("M. m. domesticus","M. m. musculus", "M.m.molossinus"),
           fontface = 'italic', hjust = 0.5, size=3.5)

```{r Strain.fig.merge.mouse.level, echo=FALSE}
fig1_C_v1  <- ggdraw() +
  draw_plot(Fig1_C_strain.female, x=0, y = 0, width=.5, height=.97) +
  draw_plot(Fig1_C_strain.male, x=.45, y=0, width=.5, height=.97) +

  draw_plot_label(label = c("A", "B"), size = 15,
                  x = c(0, .5), y = c(1, 1) )
fig1_C_v1
```

Error bars above are 2 S.E. of mouse average COs,

S.E. = sd( mouse average COs ) / Nmice

MOLF and SKIVE female data point is missing because there is only 1 female, so NAs were generated.
Mol text should be moved back above the points.
